---
title: "portfolio 4 stuff"
author: "Nanna Høgsholt"
date: "11/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, MuMIn) 
```



Analysis 1:
1. calculate suicide rate.
2. model age as predictor for suicide rate using aov()
3. check normality of residuals - plot(model)
4. transform the rate.
5. model transformer sui_rate

```{r}
df <- read.csv("WHO_suicide_statistics.csv")
df$sui_rate <- df$suicides_no/df$population

# analysis 1

model <- aov(sui_rate ~ age, data = df)
plot(model)

df$log_sui_rate <- log(df$sui_rate) # problems with zeros 
df$sqrt_sui_rate <- sqrt(df$sui_rate)
df$rec_sui_rate <- 1/df$sui_rate # maybe delete this because it is weird to devide by zero

modelsqrt <- aov(sqrt_sui_rate ~ age, data = df)
plot(model)

```



```{r}
# analysis 1 - transformation

model_transformed <- aov(sqrt_sui_rate ~ age, data = df)
plot(model_transformed)
plot(model_transformed, 4)
summary(model_transformed)

# What is the R^2 for age as a predictor?
# R^2 = SSM/SST
# SSM = 0.4154
# SSR = 1.5566
# SST = SSM + SSR = 1.972
# R^2 = 0.4154/1.972 = 0.21 roughly ;-)

summary(lm(formula = sqrt_sui_rate ~ age, data = df))

# and true enough, the adjusted R-squared is 0.2105 ! (21%)


ggplot(df, aes(x = age, y = sqrt_sui_rate, fill=age))+
  geom_bar(stat = "summary", fun.y=mean)
```

```{r}

df$age2 <- as.numeric(df$age)


ggplot(df, aes(x = age, y = suicides_no, color = age))+
  geom_bar(stat = "summary", fun.y=mean, position = "dodge", fill = "#e9e9e9")+
  ggtitle("Suicide numbers per age")

ggplot(df, aes(x = age2, y = suicides_no, color = age2))+
  geom_point(stat = "summary", fun.y=mean)+ geom_smooth(method=lm, color="red")+
  ggtitle("Suicide numbers per age")

ggplot(df, aes(x = age2, y = (sqrt_sui_rate), color = age2))+
  geom_point(stat = "summary", fun.y=mean)+ geom_smooth(method=lm, color="red")+
  ggtitle("Suicide numbers per age")

ggplot(df, aes(x = age2, y = sqrt_sui_rate))+
  geom_point(stat = "summary", fun.y=mean)+ geom_smooth(method=lm, color="#3293e3")+
  ggtitle("Suicide rate per age")+ xlab("Age groups (1=05-14y, 2=15-24, 3=25-34, 4=35-54, 5=55-74, 6=75+)")

ggplot(df, aes(x = age2, y = (sqrt_sui_rate*100), color = age2))+
  geom_point()+ geom_smooth(method=lm, color="#3293e3")+
  ggtitle("Suicide rate per age")+ xlab("Age groups (1=05-14y, 2=15-24, 3=25-34, 4=35-54, 5=55-74, 6=75+)") + ylab("sqrt(Suicide rate)*100")


ggplot(df, aes(x = age, y = sui_rate))+
  geom_point() + geom_smooth(model=lm)+
  ggtitle("Suicide numbers per year")

```

Boxplot
```{r}
ggplot(df, aes(x = age, y = sqrt_sui_rate, color=age)) +
  geom_boxplot(width = 0.5) +
  ggtitle("Age in suicide rates")+
  stat_summary(fun.y = mean, geom = "point", shape = 23, colour = "Black")
```






Analysis 2:
Suicide rates as predicted by year. (suicide rates numeric, year integer - does this meet assumptions, roughly? haha)
1. model age as predictor for sqrt_sui_rate using aov()
2. check normality of residuals - plot(model)
3. summary(lm()) for effect size / R^2

```{r}
a2model <- aov(sqrt_sui_rate ~ year, data=df)
plot(a2model) # tells us about residuals, looks way more homoschedastic than that of analysis 1 :-)
summary(a2model) # significant P-value
summary(lm(formula = sqrt_sui_rate ~ year, data=df)) # R^" = 0.0039 (meaning this model accounts for about 0.4% af variance)
```

let's visualize it a sec
```{r}
ggplot(df, aes(x = year, y = suicides_no, color = age))+
  geom_bar(stat = "summary", fun.y=mean, position = "dodge", fill = "#e9e9e9")+
  ggtitle("Suicide numbers per year")

#This is pretty but not rate, so don't use this
ggplot(df, aes(x = year, y = suicides_no, color = age))+
  geom_point(stat = "summary", fun.y=mean)+ geom_smooth(method=lm)+
  ggtitle("Suicide numbers per year")

ggplot(df, aes(x = year, y = suicides_no))+
  geom_point(stat = "summary", fun.y=mean) + geom_smooth(method=lm)+
  ggtitle("Suicide numbers per year")

ggplot(df, aes(x = year, y = sui_rate, color = age))+
  geom_bar(stat = "summary", fun.y=mean, position = "dodge", fill = "#e9e9e9")+
  ggtitle("Suicide rates per year")

#This one is good
ggplot(df, aes(x = year, y = (sqrt_sui_rate*100), color = age))+
  geom_point(stat = "summary", fun.y=mean)+ geom_smooth(method = "lm")  +
  ggtitle("Suicide rates per year") + ylab("sqrt(Suicide rate)*100")

ggplot(df, aes(x = year, y = sqrt_sui_rate))+
  geom_point(stat = "summary", fun.y=mean, position = "dodge", fill = "#e9e9e9")+ geom_smooth(method=lm)+
  ggtitle("Suicide rates per year")

```

```{r}
# just exploring

# suicides_no :
# country accounts for 41.6%
# year accounts for 0.01%
# sex accounts for 1.6%
# age accounts for 2.6%
# population accounts for 37.3%
# that is roughly a total of 54.7% of the varience explained by theoretically justifiable variables measured.


# sqrt_sui_rate :
# country accounts for 29.5%
# year accounts for 0.4%
# sex accounts for 15.4%
# age accounts for 21%
# that is roughly a total of 66.3% of the varience explained by theoretically justifiable variables measured. (rate is already dependent on suicides_no and population, so that would violate the collinearity assumption)


# sui_rate :
# country accounts for 22.4%
# year accounts for 0.5%
# sex accounts for 14%
# age accounts for 14%
# that is roughly a total of 51% of the varience explained by theoretically justifiable variables measured. (rate is already dependent on suicides_no and population, so that would violate the collinearity assumption)

summary(lm(formula = suicides_no ~ population, data=df))

```





Analysis 3:
lm(sui_rate (or sqrt_sui_rate?) ~ relevant prdictors

```{r}
# analysis 3
# model comparison

# We have 4 theoretically justifiable predictors in our maximum model (country, year, sex, age)
# 1x4. 4x3. 6x2. 4x1. = 15 different models possible.

maximal_model <- lm(formula = sqrt_sui_rate ~ country + year + sex + age, data=df)

n_country_model <- lm(formula = sqrt_sui_rate ~ year + sex + age, data=df)
n_year_model <- lm(formula = sqrt_sui_rate ~ country + sex + age, data=df)
n_sex_model <- lm(formula = sqrt_sui_rate ~ country + year + age, data=df)
n_age_model <- lm(formula = sqrt_sui_rate ~ country + year + sex, data=df)

cy_model <- lm(formula = sqrt_sui_rate ~ country + year, data=df)
cs_model <- lm(formula = sqrt_sui_rate ~ country + sex, data=df)
ca_model <- lm(formula = sqrt_sui_rate ~ country + age, data=df)
ys_model <- lm(formula = sqrt_sui_rate ~ year + sex, data=df)
ya_model <- lm(formula = sqrt_sui_rate ~ year + age, data=df)
sa_model <- lm(formula = sqrt_sui_rate ~ sex + age, data=df)

summary(maximal_model) # maximal model, R^2 = 66.28%

summary(n_country_model) # R^2 = 36.86%
summary(n_year_model) # R^2 = 66.14%
summary(n_sex_model) # R^2 = 50.81%
summary(n_age_model) # R^2 = 45.15%

summary(cy_model) # R^2 = 29.69%
summary(cs_model) # R^2 = 45.01%
summary(ca_model) # R^2 = 50.67%
summary(ys_model) # R^2 = 15.81%
summary(ya_model) # R^2 = 21.45%
summary(sa_model) # R^2 = 36.47%

```



```{r}
ggplot(df, aes(x = sex, y = sui_rate, color = age))+
  geom_bar(stat = "summary", fun.y=mean, position = "dodge", fill = "#e9e9e9", size = 1.2)
```

