---
title: "Portfolio 5"
author: "Nanna Høgsholt"
date: "29/11/2019"
output: html_document
---

*Portfolio 5 
```{r}
library(pacman);p_load(tidyverse,lme4, boot, lmerTest, caret, e1071,ggplot2)
```

**Analysis 1

```{r}
#Loading data
cake<-read.csv("cake.csv", header = T)

```

Boxplotting
```{r}
boxplot(angle ~ recipe, col=c("white","lightgrey"),cake)
boxplot(angle ~ temp, col=c("white","lightgrey"),cake)
boxplot(angle ~ recipe+temp, col=c("white","lightgrey"),cake)
boxplot(angle ~ temp+recipe, col=c("white","lightgrey"),cake)
boxplot(angle ~ replicate, col=c("white","lightgrey"),cake)
boxplot(angle ~ recipe + temp, cake)
boxplot(angle ~ replicate + temp, cake)

```

```{r}
#adding random intercepts
cake.model=lmerTest::lmer(angle ~ temp+(1|recipe)+(1|replicate), data=cake, REML=F)
summary(cake.model)

```

variance of angle breage of each recipe 

Intercept model
```{r}
m1 <- lmerTest::lmer(angle ~ recipe + (1|replicate),data = cake, REML = F)
m2 <- lmerTest::lmer(angle ~ recipe + temp + (1|replicate),data = cake, REML = F)
m3 <- lmerTest::lmer(angle ~ temp + (1|recipe),data = cake, REML = F)
m4 <- lmerTest::lmer(angle ~ recipe * temp + (1|replicate), cake, REML= FALSE)
m5 <- lmerTest::lmer(angle ~ recipe + (1+temp|recipe),data = cake, REML = F)
m6 <- lmerTest::lmer(angle ~ recipe + (1+replicate|recipe),data = cake, REML = F)
m7 <- lmerTest::lmer(angle ~ recipe + (1+replicate|temp),data = cake, REML = F)
m8 <- lmerTest::lmer(angle ~ recipe + temp + (1|recipe),data = cake, REML = F)
m9 <- lmerTest::lmer(angle ~ temp + replicate + (1|recipe),data = cake, REML = F)
m10 <- lmerTest::lmer(angle ~ temp + (1+temp|recipe),data = cake, REML = F)
m11 <- lmerTest::lmer(angle ~ recipe + temp + (1+temp|recipe),data = cake, REML = F)
m12 <- lmerTest::lmer(angle ~ recipe + (1|temp),data = cake, REML = F)
m13 <- lmerTest::lmer(angle ~ temp + recipe + (1+temp|recipe),data = cake, REML = F)
m14 <- lmerTest::lmer(angle ~ recipe + temp + (1+recipe|replicate),data = cake, REML = F)
m15 <- lmerTest::lmer(angle ~ recipe + (1+replicate|temp),data = cake, REML = F)
m16 <- lmerTest::lmer(angle ~ recipe + temp+ (1+replicate|temp),data = cake, REML = F)
#m17 makes hella sense RN. So on the y axis is angle and on the x axis is temperature. Because for each replicate of the recipe is a corresponding temperature we can give replicate indedependent intercept and recipes independent slopes.
m_1 <- lmerTest::lmer(angle ~ temp +  (1+recipe|replicate),data = cake, REML = F)
m_2 <- lmerTest::lmer(angle ~ temp +  (1+replicate|recipe),data = cake, REML = F)
#It makes sence to add recipe as a fixed variable because we expect different outcomes from different cake recipies 
m_3 <- lmerTest::lmer(angle ~ temp + recipe+ (1+recipe|replicate),data = cake, REML = F)
#m20 recipe can  be fixed because we expect different outcomes from each recipe 
m20 <- lmerTest::lmer(angle ~ temp +  recipe + (1+replicate|recipe),data = cake, REML = F)
m21 <- lmerTest::lmer(angle ~ temp + (1+recipe|recipe),data = cake, REML = F)
m22 <- lmerTest::lmer(angle ~ temp +  recipe + (1|replicate),data = cake, REML = F)

anova(m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11, m12, m13, m14, m15, m16, m17, m18, m19, m20, m21, m22)


summary(intercept_model)
summary(intercept_model2)
summary(m3)
summary(m4)
summary(m5)

```
The best models
```{r}
m17 <- lmerTest::lmer(angle ~ temp +  (1+recipe|replicate),data = cake, REML = F)
m18 <- lmerTest::lmer(angle ~ temp +  (1+replicate|recipe),data = cake, REML = F)
#It makes sence to add recipe as a fixed variable because we expect different outcomes from different cake recipies 
m19 <- lmerTest::lmer(angle ~ temp + recipe+ (1+recipe|replicate),data = cake, REML = F)

#Creating new models after Fabio
m01<-lmerTest::lmer((angle~recipe + (1|replicate)),data = cake, REML = F)
#failed: m02<-lmerTest::lmer((angle~recipe + (1+recipe|replicate)),data = cake, REML = F)
m03<-lmerTest::lmer((angle~temp + (1|replicate)),data = cake, REML = F)
#failed: m04<-lmerTest::lmer((angle~temp + (1+temp|replicate)),data = cake, REML = F)
#boundary: m05<-lmerTest::lmer((angle~recipe + temp+ (1+recipe|replicate)),data = cake, REML = F)
m06<-lmerTest::lmer((angle~recipe + temp + (1|replicate)),data = cake, REML = F)

#try these
	- (angle~recipe + (1|replicate))
	- (angle~recipe+(1+reicpe´|replicate)
	- (angle~temp + (1|replicate))
	- (angle~temp+(1+temp´|replicate)
  - (angle~ temp + recipe+(1+reicpe´|replicate)

anova(m01, m03, m06)
```
Loglikelyhood like SSR how much uexplain vairance, the larger the likelyhood the more unexplained variables. 



Chosing the best model fra AIC is m17, where angle is predicted by temperature, recipe has independent slopes and replicate has independent intercepts

```{r}
summary(m06)
#R squared
MuMIn::r.squaredGLMM(m06)
```



```{r}
ggplot(cake, aes(x=temp, y=angle, color=replicate, shape=recipe)) + geom_point() + geom_smooth(method = lm)

```
Maybe subsetting is a better visual?
```{r}
cakeA<-filter(cake, recipe=="A")
cakeB<-filter(cake, recipe=="B")
cakeC<-filter(cake, recipe=="C")
```

```{r}
ggplot(cakeA, aes(x=temp, y=angle, color=replicate)) + geom_point() + geom_smooth(method = lm) + ggtitle("Recipe A")
ggplot(cakeB, aes(x=temp, y=angle, color=replicate)) + geom_point() + geom_smooth(method = lm) + ggtitle("Recipe B")
ggplot(cakeC, aes(x=temp, y=angle, color=replicate)) + geom_point() + geom_smooth(method = lm) + ggtitle("Recipe C")
```
These plots prove that we need independent slopes and independent intercepts to make de best model

```{r}
library(rgl)
plot3d(cake$recipe, cake$angle, cake$temp,size=1, type="s", col = cake$recipe)
```



```{r}
coef(m17)
```



```{r}
plot(m06)
```

```{r}
qqplot(m17)
```


GGPlot
```{r}
ggplot(cake, aes(x=replicate,y=angle, color=recipe)) +
  geom_point() + geom_smooth(aes(model="lm"))
```



Make a max model
```{r}
max.mod<-lmerTest::lmer(angle ~ recipe + temp + replicate(1|),data = cake, REML = F)
```

